---
title: "TEXT ANALYTICS"
author: "Bharath"
date: '2022-04-19'
output:
  pdf_document: default
  html_document: default
---

```{r}
pacman::p_load(tidyverse, tidytext, lubridate, stringr, reshape2, wordcloud, textdata)
```

```{r}
load(file='song_lyrics.rda')
song_lyrics
```
1. Tokenize the lyrics variable, count the number of words in each song, and merge the
total number of words in each song with the data (this will add a new column to the
data). Print a few rows of the updated data set showing the total number of words in each
song.



```{r}
for( i in 1:length(song_lyrics$lyrics)){
  text_df <- tibble( text = song_lyrics$lyrics[i])
  k=text_df %>% unnest_tokens(word, text)
  
  song_lyrics$count[i] = unlist(count(k))
  
}
song_lyrics$count = unlist(song_lyrics$count)
song_lyrics_count = song_lyrics[c("song","count")]
head(song_lyrics)
head(song_lyrics_count)
```




```{r}
text_df <- tibble( text = song_lyrics$lyrics)

text_df =text_df %>% unnest_tokens(word, text)
```



```{r}
data("stop_words")
stop_words
filter(stop_words, lexicon == "bing")

```



```{r}
text_df %>%
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  top_n(20) %>% 
  ggplot(aes(fct_reorder(word, n), n)) + 
  geom_col(fill="gold3", width = 0.6) +
  coord_flip() +
  labs(y="Frequency", x = NULL, title = "Most Frequent Words- Top 20")

```
```{r}
text_df %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word, sort = TRUE) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  ggplot(aes(fct_reorder(word, n), 
             n, 
             fill = sentiment)) +
  geom_col(width = 0.4) +
  coord_flip() +
  facet_wrap(~ sentiment, scales = "free") +
  labs(y= "Contribution to Sentiment", x = NULL)
```
```{r}
text_df %>% # total no of positive and negative
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, sort = TRUE)
```


```{r}
m <- song_lyrics %>%
count(song) %>%
rename(words = n)
m
```
```{r}
song_lyrics_1 <- unnest_tokens(song_lyrics, word, lyrics) %>%
left_join(m, by = 'song') ## left join to get all the song lyrics matched
song_lyrics_1
```

```{r}
song_lyrics_2 = inner_join(song_lyrics_1, get_sentiments("bing"))
song_lyrics_2
```


-> Create sentiment scores using “bing” lexicon and count the number of sentiment words
(number of positive and number of negative sentiment words, separately) in each song.
Print two separate tables: (i) songs with the highest percentage of positive words; (ii)
songs with the highest percentage of negative words.



```{r}
Neg_sen <- song_lyrics_2 %>%
count(song, sentiment, count) %>%
ungroup() %>%
filter(sentiment == "negative") %>%
mutate(percent = round((n / count)*100,2)) %>%
rename("Sentiment count" = n)
Neg_sen = Neg_sen[c("song","percent","sentiment")][order(-Neg_sen$percent),]

```


```{r}

pos_sen <- song_lyrics_2 %>%
count(song, sentiment, count) %>%
ungroup() %>%
filter(sentiment == "positive") %>%
mutate(percent = round(  (n / count)*100,2)) %>%
rename("Sentiment count" = n)
pos_sen = pos_sen[c("song","percent","sentiment")][order(-pos_sen$percent),]
```

```{r}
Neg_sen
pos_sen

```
-> The songs "shake it off" and "shake shake shake shake your booty" has highest negative words percentage with (47.3% && 45.63%) and followed by "disturbia" ans so on.
 The songs "diamonds" has highest positive word rate with (57.97%)


3. Do songs with a higher percent of negative words rank higher on the Billboard chart
(rank 1 is the highest and rank 100 is the lowest)? Answer this question by grouping
songs in the following groups: 1-10, 11-20, ..., and 91-100. Use boxplots for each of these
groups, with ranking groups on the x axis and percent of negative words on the y axis.

```{r}
rank <- song_lyrics_2 %>%
filter(sentiment == "negative") %>%
count(song, rank,count) %>%
ungroup() %>%
rename(negative_count = n)
```


```{r}
rank1 = rank %>%
mutate(percent = (negative_count / count)*100, rank = 10 * floor(rank / 10))
```


```{r}
rank2 = rank1[order(-rank1$percent),]

```

```{r}
ggplot(distinct(rank2), aes(as.factor(rank), percent)) +
geom_boxplot()+
xlab("Rank[1-100]")+
ylab("negative words percentage")+
ggtitle("Rank analysis")
```
-> There is a low percent of negative words at rank 1-10 and and also at 90-100.also there are songs with high negative words at top ranks. This tells us that negative words cannot describe the rank.


4. Have the percent of positive words in a song increased over the decade? Answer this
question by grouping songs over decades: 1960-1969, 1970-1979, ..., 2010-2019. Use
boxplots for each of these groups, with decades on the x axis and the percent of positive
words on the y-axis.


```{r}
year <- song_lyrics_2 %>%
filter(sentiment == "positive") %>%
count(song, year,count) %>%
ungroup() %>%
rename(negative_count = n) 


year1 = year %>%
mutate(percent = (negative_count / count)*100, year = 10 * floor(year / 10))


year2 = year1[order(-year1$percent),]
```



```{r}
ggplot(distinct(year1), aes(as.factor(year), percent)) +
geom_boxplot()+
xlab("years")+
ylab("positive words percentage")+
ggtitle("Year analysis")
```
-> on observing the above boxplot the positive trend is located from 1960 to 1970. Then after there is a decreasing trend is observed from 1970 to 2000 years. Then a small sloped positive trend is observed till 2010. On average  a decreasing trend is observed.
